---
title: "SRA Workshop 7"
author: "John Benninghoff"
date: '2021-12-05'
file-modified: '2022-02-06'
categories: []
order:
output:
  html_notebook:
    theme:
      version: 5
      preset: bootstrap
    css: assets/extra.css
    pandoc_args: --shift-heading-level-by=1
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
---

Working notebook for SRA 2021 Workshop 7, "Monte Carlo simulation and probability bounds analysis in R with hardly any data (Instructors: Ferson & Grey)"

```{r setup, message=FALSE, warning=FALSE}
library(MASS)
library(sn)
```

# Instructions

Download `sra.r` and `pba.r` from <https://sites.google.com/site/hardlyanydata> and source it into R. Set `RStudio = TRUE` for use within RStudio and save the file. The library also requires the package `sn`. Use `rm(list = ls())` to clear R environment.

There is another version of `pba.r` on [GitHub](https://github.com/ScottFerson/pba.r)

# Monte Carlo

Monte Carlo example from class using `sra.R`; contaminant plume (from Lobascio), slide 37-39 in the workshop [PowerPoint](assets/NoData-2021-Zoom.pptx):

View the cumulative probability plot by entering the letter in the console.

```{r T}
source("assets/sra.R")

# nolint start
L <- uniform(80, 120) # [m], source-receptor distance
i <- uniform(0.0003, 0.0008) # [], hydraulic gradient
K <- lognormal(1000, 750) # [m yr1], hydraulic conductivity
n <- lognormal(0.25, 0.05) # [], effective soil porosity
BD <- lognormal(1650, 100) # [kg per m3], soil bulk density
foc <- uniform(0.0001, 0.005) # fraction organic carbon
Koc <- normal(10, 3) # [m3 per kg], partition coefficient

T <- (n + BD * foc * Koc) * L / (K * i) # all variables assumed independent
summary(T)
T
# nolint end
```

Truncated version, from slide 40:

```{r T_truncated}
# nolint start
L <- uniform(80, 120) # [m], source-receptor distance
i <- uniform(0.0003, 0.0008) # [], hydraulic gradient
K <- lognormal(1000, 750) # [m yr1], hydraulic conductivity
K <- truncate(K, 300, 3000)
n <- lognormal(0.25, 0.05) # [], effective soil porosity
n <- truncate(n, 0.2, 0.35)
BD <- lognormal(1650, 100) # [kg per m3], soil bulk density
BD <- truncate(BD, 1500, 1750)
foc <- uniform(0.0001, 0.005) # fraction organic carbon
Koc <- normal(10, 3) # [m3 per kg], partition coefficient
Koc <- truncate(Koc, 5, 20)

T <- (n + BD * foc * Koc) * L / (K * i)
summary(T)
T
# nolint end
```

Maximum entropy version, from slide 70:

T and Tind are pretty similar. In this case, the maximum entropy approach comes up with a similar distribution as the truncated approach.

```{r Tind}
# nolint start
L <- MEmmms(80, 120, 100, 11.55) # source-receptor distance
i <- MEmmms(0.0003, 0.0008, 0.00055, 0.0001443) # hydraulic gradient
K <- MEmmms(300, 3000, 1000, 750) # hydraulic conductivity
n <- MEmmms(0.2, 0.35, 0.25, 0.05) # effective soil porosity
BD <- MEmmms(1500, 1750, 1650, 100) # soil bulk density
foc <- MEmmms(0.0001, 0.005, 0.00255, 0.001415) # fraction organic carbon
Koc <- MEmmms(5, 20, 10, 3) # organic partition coefficient

Tind <- (n + BD * foc * Koc) * L / (K * i)
summary(Tind)
Tind
# nolint end
```

# Probability Bounds Analysis

Conduct the same analysis using PBA, from slide 162. See the important note from slide 162 below.

*Note: N.B. The implementation of mmms in pba.r is incomplete so, while its results are bounds, they are not best possible bounds*

```{r Tind_pba}
# ideally we'd clear the environment with `rm(list = ls())`, but that doesn't work here;
# an alternative would be to move this to a separate notebook. This approach appears to be OK,
# outputs from running just the setup and the entire notebook are the same.

# this code breaks build_analysis_site() - commented out because of that.
# uncomment this code and it should work within the notebook in RStudio.
# nolint start
# source("assets/pba.R")
#
# L <- mmms(80, 120, 100, 11.55) # source-receptor distance
# i <- mmms(0.0003, 0.0008, 0.00055, 0.0001443) # hydraulic gradient
# K <- mmms(300, 3000, 1000, 750) # hydraulic conductivity
# n <- mmms(0.2, 0.35, 0.25, 0.05) # effective soil porosity
# BD <- mmms(1500, 1750, 1650, 100) # soil bulk density
# foc <- mmms(0.0001, 0.005, 0.00255, 0.001415) # fraction organic carbon
# Koc <- mmms(5, 20, 10, 3) # organic partition coefficient
#
# up <- 100000 # detail
# Tind <- (n + BD * foc * Koc) * L / (K * i)
# Tind <- pmin(Tind, up)
#
# summary(Tind)
# Tind
# nolint end
```
